Weak interspecific interactions in a sagebrush steppe: Evidence from observational data, population models, and a removal experiment

### 2016-09-20 removals/manuscript/removal_ms.tex
Revising manuscript

\begin{enumerate}
\item A previous study of four dominant species from a sagebrush steppe community used historical data sets and multispecies population models to demonstrate weak interspecific interactions, implying very little competitive release. 
\item We conducted a removal experiment to test the model's ability to predict competitive release. We established new quadrats in the same location where the historical data were collected. We assigned half the new quadrats to a  perennial grass removal treatment and half to a treatment in which we removed the dominant shrub, \textit{Artemisia tripartita}.
\item We modeled survival, growth and recruitment as functions of local neighborhood species composition along with an indicator variable for removal treatment. For survival and recruitment, the removal treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the ``baseline'' model. 
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models  based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected much stronger competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that \new{interspecific} resource competition is not the primary factor determining the dynamics of this community.

@@

An alternative approach to studying competitive release is to fit multispecies population models to long-term, observational data, and then simulate population responses to removals. Applying this approach to arid and semiarid plant communities, we found evidence of strong intraspecific limitation and very weak interspecific interactions, leading to very stable coexistence and implying very little competitive release \citep{Adler:2010fu} ( Chu and Adler 2015). Since we only studied the most common, co-occurring species, we expected to find evidence of stable coexistence, but we were surprised that interspecific interactions and species' responses to the loss of a dominant competitor were so weak. Compared to traditional field experiments, our approach has the advantage of projecting full life-cycle responses over long time periods, but its reliance on observational data sets limits the strength of our inference. We would have more confidence in our results if they were confirmed by an experiment.

@@

% add climate data for 2011-2015

@@

@@

@@

@@

Trends in cover by species and removal treatment were variable. For example, grass removal had little effect on the cover of \textit{A. tripartita} 
(Fig. \ref{fig:CoverTrends}), but shrub removal did appear to cause an increase in the cover of \textit{P. spicata}. We tested the effects of removal on cover by analyzing year-to-year changes in (log) cover, a measure of population growth rate (Fig. \ref{fig:CoverChange}). The removal treatments had no significant effect on 
\textit{A. tripartita} (95\% confidence interval on removal treatment effect: -0.45, 0.27) or 
\textit{Poa secunda} (95\% CI: -0.40, 0.28) population growth, but did increase 
\textit{H. comata} (95\% CI: 0.02, 0.87) and
 \textit{P. spicata} (95\% CI: 0.04, 0.50) population growth 
(Table \ref{table:coefficients}). However, these treatment comparisons do not address our questions about whether responses to removal were 
adequately predicted by models fit to observational data on non-removal plots.

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=1\textwidth]{cover_change_chrono}
\caption{Observed and predicted year-to-year changes in cover for the four dominant species. Mean observed changes (averaged across quadrats) are shown in solid symbols, mean predicted changes are shown in open symbols. Values for control and removal plots are blue and red, respectively. Predictions from the ``Baseline'' model do account for effects of local species composition on vital rates, but do not include additional effects of removals. Predictions from the ``Removal'' model do include additional treatment effects on vital rates. }
\label{fig:CoverChange}
\end{figure}

@@

Our vital rate models take into account the effects of local species composition and should therefore be able to predict the effect of removals on survival, growth and recruitment without any additional parameters. If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of removals that goes above and beyond the response predicted by a model based on data from the control plots. For the survival models, the 95\% confidence intervals on the removals effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data from control plots. However, for the growth models, the 95\% confidence intervals on the removal effects were positive for \textit{Poa secunda}  and \textit{P. spicata}, meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). In contrast, \textit{Poa secunda} had lower recruitment in removal than controls plots (upper limit of 95\% CI $<0$; Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}).

@@

Second, we evaluated the possibility that removal effects might be transient. We removed aboveground plant material, but roots of the removed plants remain in the ground and decompose, potentially creating a flush of mineral nutrients (refs?). We tested for a transient removal effect by adding a year-by-treatment interaction. For both  \textit{Poa secunda}  and \textit{P. spicata} , we found no clear sign of a decrease in the removal effect over time, and in fact the removal effects for \textit{P. spicata} were strongest in the last two years of observations (Tables \ref{table:POSEgrowth-byYr} and \ref{table:PSSPgrowth-byYr}).

@@

To quantify the short-term population-level consequences of removals, we used an IBM to project the cover of each species in each quadrat one year ahead using versions of the model with and without removal treatment effects. For \textit{A. tripartita}, both observed and predicted population growth rates were close to zero, and differences in model predictions were small (Fig. \ref{fig:CoverChange}). \textit{H. comata} experienced greater variation in growth rates, but the inclusion of removal treatment effects had little effect on population growth rates. For \textit{Poa secunda} and especially \textit{P. spicata}, the model with removal effects consistently predicted higher population growth than the baseline model  (Fig. \ref{fig:CoverChange}). For \textit{P. spicata}, including the removal effects increased the annual multiplicative rate of increase by roughly 35\%.

To quantify the long-term population-level consequences of removals, we used an IPM to project stochastic equilibrium cover. We concentrated on grass species' responses to \textit{A. tripartita} removal, since the growth regressions indicated our baseline model might underestimate competitive release in this situation. Using the baseline model, with no removal treatment coeffiicients, the IPM predicts little effect of \textit{A. tripartita} removal on the cover of any grass species (compare gray and brown boxes in Fig. \ref{fig:IPMresults}). Using a model with removal treatment coefficients, the IPM predicts little effect of shrub removal on \textit{Poa secunda} and \textit{H. comata}, but a 2.75 fold increase in the cover of \textit{P. spicata} (Fig. \ref{fig:IPMresults}).

@@

\begin{table}[h]
\caption{Cover change models}
\centering
%\begin{center}
\begin{tabular}{l c c c c }
\hline
Species & ARTR & HECO & POSE & PSSP \\
\hline
(Intercept)           & $-0.03$          & $-0.04$          & $-0.02$          & $-0.09$          \\
                      & $[-0.33;\ 0.27]$ & $[-0.32;\ 0.24]$ & $[-0.31;\ 0.27]$ & $[-0.33;\ 0.15]$ \\
Treatment: No grasses    & $-0.09$          &                  &                  &                  \\
                      & $[-0.45;\ 0.27]$ &                  &                  &                  \\
Treatment: No shrub    &                  & $0.44^{*}$       & $-0.06$          & $0.27^{*}$       \\
                      &                  & $[0.02;\ 0.87]$  & $[-0.40;\ 0.28]$ & $[0.04;\ 0.50]$  \\
\hline
AIC                   & 23.20            & 59.58            & 162.06           & 117.50           \\
BIC                   & 33.15            & 66.24            & 173.51           & 129.22           \\
Log Likelihood        & -6.60            & -24.79           & -76.03           & -53.75           \\
Num. obs.             & 54               & 28               & 73               & 77               \\
Num. groups: quad     & 12               & 5                & 14               & 14               \\
Num. groups: year     & 8                & 8                & 8                & 8                \\
Var: quad (Intercept) & 0.08             & 0.00             & 0.00             & 0.00             \\
Var: year (Intercept) & 0.01             & 0.00             & 0.10             & 0.09             \\
Var: Residual         & 0.04             & 0.32             & 0.40             & 0.19             \\
\hline
\multicolumn{5}{l}{\scriptsize{$^*$ 0 outside the confidence interval}}
\end{tabular}
\label{table:coefficients}
%\end{center}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Thu May 19 08:31:09 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the ARTR survival model} 
\label{ARTRsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & -0.0705 & 0.2256 & -0.5109 & -0.0723 & 0.3799 & -0.0757 & 0.0000 \\ 
  logarea & 0.7137 & 0.0605 & 0.5980 & 0.7124 & 0.8370 & 0.7098 & 0.0000 \\ 
  TreatmentNo\_grass & -0.2581 & 0.8998 & -1.9434 & -0.2884 & 1.5959 & -0.3511 & 0.0000 \\ 
  W.ARTR & -1.4362 & 0.2154 & -1.8709 & -1.4322 & -1.0242 & -1.4240 & 0.0000 \\ 
  W.HECO & -0.0818 & 0.0430 & -0.1630 & -0.0830 & 0.0062 & -0.0855 & 0.0000 \\ 
  W.POSE & -0.1159 & 0.0859 & -0.2851 & -0.1158 & 0.0523 & -0.1155 & 0.0000 \\ 
  W.PSSP & 0.0292 & 0.0276 & -0.0225 & 0.0283 & 0.0863 & 0.0263 & 0.0000 \\ 
  W.allcov & -0.0081 & 0.0093 & -0.0265 & -0.0081 & 0.0099 & -0.0080 & 0.0000 \\ 
  W.allpts & 0.1296 & 0.1579 & -0.1787 & 0.1290 & 0.4409 & 0.1278 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Thu May 19 08:31:59 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the HECO survival model} 
\label{HECOsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.4652 & 0.2285 & 1.0126 & 1.4652 & 1.9170 & 1.4650 & 0.0000 \\ 
  logarea & 1.3190 & 0.1036 & 1.1266 & 1.3149 & 1.5346 & 1.3068 & 0.0000 \\ 
  TreatmentNo\_shrub & 0.7313 & 0.4426 & -0.1292 & 0.7284 & 1.6074 & 0.7226 & 0.0000 \\ 
  W.ARTR & -0.0049 & 0.0037 & -0.0119 & -0.0050 & 0.0027 & -0.0052 & 0.0000 \\ 
  W.HECO & -0.6707 & 0.0725 & -0.8206 & -0.6680 & -0.5354 & -0.6627 & 0.0000 \\ 
  W.POSE & 0.0965 & 0.0690 & -0.0272 & 0.0921 & 0.2440 & 0.0830 & 0.0000 \\ 
  W.PSSP & -0.0015 & 0.0197 & -0.0386 & -0.0020 & 0.0390 & -0.0032 & 0.0000 \\ 
  W.allcov & -0.0032 & 0.0058 & -0.0140 & -0.0034 & 0.0087 & -0.0038 & 0.0000 \\ 
  W.allpts & -0.2630 & 0.1184 & -0.4889 & -0.2654 & -0.0236 & -0.2701 & 0.0000 \\ 
  logarea:W.ARTR & -0.0005 & 0.0028 & -0.0057 & -0.0006 & 0.0052 & -0.0008 & 0.0000 \\ 
  logarea:W.HECO & 0.2751 & 0.0600 & 0.1522 & 0.2770 & 0.3879 & 0.2809 & 0.0000 \\ 
  logarea:W.POSE & 0.0067 & 0.0513 & -0.0833 & 0.0026 & 0.1182 & -0.0060 & 0.0000 \\ 
  logarea:W.PSSP & -0.0273 & 0.0147 & -0.0543 & -0.0280 & 0.0036 & -0.0294 & 0.0000 \\ 
  logarea:W.allcov & -0.0043 & 0.0041 & -0.0118 & -0.0045 & 0.0044 & -0.0050 & 0.0000 \\ 
  logarea:W.allpts & -0.1474 & 0.0903 & -0.3187 & -0.1496 & 0.0364 & -0.1541 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Thu May 19 08:33:14 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the POSE survival model} 
\label{POSEsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.3190 & 0.2047 & 0.9110 & 1.3199 & 1.7215 & 1.3213 & 0.0000 \\ 
  logarea & 1.0711 & 0.0694 & 0.9387 & 1.0693 & 1.2137 & 1.0657 & 0.0000 \\ 
  TreatmentNo\_shrub & -0.0442 & 0.1832 & -0.4030 & -0.0445 & 0.3158 & -0.0451 & 0.0000 \\ 
  W.ARTR & 0.0001 & 0.0019 & -0.0036 & 0.0001 & 0.0038 & 0.0001 & 0.0000 \\ 
  W.HECO & -0.0075 & 0.0142 & -0.0353 & -0.0076 & 0.0204 & -0.0077 & 0.0000 \\ 
  W.POSE & -1.2650 & 0.0825 & -1.4296 & -1.2641 & -1.1053 & -1.2624 & 0.0000 \\ 
  W.PSSP & 0.0244 & 0.0123 & 0.0005 & 0.0243 & 0.0486 & 0.0242 & 0.0000 \\ 
  W.allcov & -0.0031 & 0.0027 & -0.0083 & -0.0031 & 0.0022 & -0.0031 & 0.0000 \\ 
  W.allpts & 0.0145 & 0.0527 & -0.0889 & 0.0144 & 0.1181 & 0.0143 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

\begin{enumerate}
\item A previous study of four dominant species from a sagebrush steppe community used historical data sets and multispecies population models to demonstrate weak interspecific interactions, implying very little competitive release. 
\item We conducted a removal experiment to test the model's ability to predict competitive release. We established new quadrats in the same location where the historical data were collected. We assigned half the new quadrats to a  perennial grass removal treatment and half to a treatment in which we removed the dominant shrub, \textit{Artemisia tripartita}. We modeled survival, growth and recruitment as functions of local neighborhood species composition along with an indicator variable for removal treatment. If our ``baseline" model, which accounts for local plant-plant interactions, accurately predicts responses to removal, then the fitted removal indicator effect should be non-signficant.
\item  For survival and recruitment, the removal treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the baseline model. For  \textit{Poa secunda} recruitment, the removal treatment was actually negative, indicating that our baseline model overestimated competitive release.
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected greater competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that \new{interspecific} resource competition is not the primary factor determining the dynamics of this community.

@@

An alternative approach to studying competitive release is to fit multispecies population models to long-term, observational data, and then simulate population responses to removals. Applying this approach to arid and semiarid plant communities, we found evidence of strong intraspecific limitation and very weak interspecific interactions, leading to very stable coexistence and implying little competitive release \citep{Adler:2010fu} ( Chu and Adler 2015). Since we only studied the most common, co-occurring species, we expected to find evidence of stable coexistence, but we were surprised that interspecific interactions and species' responses to the loss of a dominant competitor were so weak. Compared to traditional field experiments, our approach has the advantage of projecting full life-cycle responses over long time periods, but its reliance on observational data sets limits the strength of our inference. We would have more confidence in our results if they were confirmed by an experiment.

@@

@@

One reason we might observe small effects of the removal treatment is the limited power of our experiment. Perhaps large uncertainty in the estimates of these effects makes it difficult to reject the null hypothesis represented by our baseline model. To address this possibility, we also simulated a version of the model in which we replaced the mean estimate of each removal treatment effect with the parameter value at either the 5\% or 95\% confidence interval, whichever was farthest from zero. This version of the model shows the population-level changes we would expect given maximum removal effects.

@@

As with the IBM, we repeated these simulations using maximum removal treatment effects represented by parameters values at the 5\% or 95\% confidence limit farthest from zero.

@@

During the period of our experiment, annual water-year (October-September) precipitation fell within the 5\% - 95\% quantiles of the long-term mean, although more years fell below the mean than above it (\ref{fig:climate}a). However, annual mean temperatures were consistently warm relative to the long-term mean, and two years were warmer than the 95\% quantile (\ref{fig:climate}b).

@@

Trends in cover by species and removal treatment were variable. For example, grass removal had little effect on the cover of \textit{A. tripartita} 
(Fig. \ref{fig:CoverTrends}), but shrub removal did appear to cause an increase in the cover of \textit{P. spicata}. 
We tested the effects of removal on cover by analyzing year-to-year changes in (log) cover, a measure of population growth rate (Fig. \ref{fig:CoverChange}). The removal treatments had no significant effect on 
\textit{A. tripartita} (95\% confidence interval on removal treatment effect: -0.75, 0.56) or 
\textit{Poa secunda} (95\% CI: -0.41, 0.25) population growth, but did increase 
\textit{H. comata} (95\% CI: 0.01, 0.65) and
 \textit{P. spicata} (95\% CI: 0.01, 0.40) population growth 
(Table \ref{table:coefficients}). However, these treatment comparisons do not address our questions about whether responses to removal were 
adequately predicted by models fit to observational data on non-removal plots.

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=1\textwidth]{cover_change_chrono}
\caption{Observed and predicted year-to-year changes in cover for the four dominant species. Mean observed changes (averaged across quadrats) are shown in solid symbols, mean predicted changes are shown in open symbols. Values for control and removal plots are blue and red, respectively. Predictions from the ``Baseline'' model do account for effects of local species composition on vital rates, but do not include additional effects of removals. Predictions from the ``Removal'' model include additional treatment effects on vital rates. }
\label{fig:CoverChange}
\end{figure}

%$  $ FIX COVERCHANGE LEGEND

@@

Our vital rate models take into account the effects of local species composition and should therefore be able to predict the effect of removals on survival, growth and recruitment without any additional parameters. If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of removals that goes above and beyond the response predicted by a model based on data from the control plots. For the survival models, the 95\% confidence intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data from control plots. However, for the growth models, the 95\% confidence intervals on the removal effects were positive for \textit{Poa secunda}  and \textit{P. spicata}, meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). In contrast, \textit{Poa secunda} had lower recruitment in removal than controls plots (upper limit of 95\% CI $<0$; Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}).

@@

Second, we evaluated the possibility that removal effects might be transient. We removed aboveground plant material, but roots of the removed plants remain in the ground and decompose, potentially creating a flush of mineral nutrients. We tested for a transient removal effect by adding a year-by-treatment interaction. For both  \textit{Poa secunda}  and \textit{P. spicata} , we found no sign of a decrease in the removal effect over time (Tables \ref{table:POSEgrowth-byYr} and \ref{table:PSSPgrowth-byYr}).

@@

To quantify the short-term population-level consequences of removals, we used an IBM to project the cover of each species in each quadrat one year ahead using versions of the model with and without removal treatment effects. Our models tended to under-predict population growth (Fig. \ref{fig:CoverChange}) and population size (Fig. \ref{fig:IBM1step}) for all species, but the difference between predictions from the baseline and removal models varied among species. For \textit{A. tripartita}, both observed and predicted population growth rates were close to zero, and differences in model predictions were small. \textit{H. comata} experienced greater variation in growth rates, but the inclusion of removal treatment effects had little effect on population growth rates. For \textit{Poa secunda} and especially \textit{P. spicata}, the model with removal effects consistently predicted higher population growth than the baseline model (Fig. \ref{fig:CoverChange}), and reduced the underestimates of cover (Fig. \ref{fig:IBM1step}). The version of the model with maximum removal treatment effects (removal coefficients set to the 95\% confidence limit furthest from zero), were qualitatively similar, but further amplified the contrast between predictions of the baseline and removal models for and \textit{P. spicata} REF MAX CI FIG

 \begin{figure}[tbp]
 \centering
 \includegraphics[width=1\textwidth]{cover_projections_1step}
 \caption{Observed and predicted cover for each of the four modeled species. Solid lines show mean observed cover averaged across control (blue) and removal treatment (red) quadrats. Dashed lines show one-step-ahead predictions from the baseline IBM (no removal treatment coefficients), averaged across quadrats. Dotted lines show one-step-ahead predictions from an IBM that includes removal treatment coefficients. }
 \label{fig:IBM1step}
 \end{figure}

% add max CI figure in SI

To quantify the long-term population-level consequences of removals, we used an IPM to project stochastic equilibrium cover. We concentrated on grass species' responses to \textit{A. tripartita} removal, since the growth regressions indicated our baseline model might underestimate competitive release in this situation (our baseline vital rate models worked well in predicting \textit{A. tripartita} responses to grass removal). Using the baseline model, with no removal treatment coeffiicients, the IPM predicts little effect of \textit{A. tripartita} removal on the cover of any grass species (compare gray and brown boxes in Fig. \ref{fig:IPMresults}). Using a model with removal treatment coefficients, the IPM predicts little effect of shrub removal on \textit{Poa secunda} and \textit{H. comata}, but a 2-fold increase in the cover of \textit{P. spicata} (Fig. \ref{fig:IPMresults}). Running a version of the model with maximum removal treatment effects (removal coefficients set to the 95\% confidence limit furthest from zero), resulted in the same qualitative outcome, but a larger increase in \textit{P. spicata} % (Fig. \ref{fig:IPMresults-maxCI})

% add max CI figure in SI

@@

\begin{table}[h]
\caption{Cover change models}
\centering
%\begin{center}
\begin{tabular}{l c c c c }
\hline
Species & ARTR & HECO & POSE & PSSP \\
\hline
(Intercept)           & $0.00$           & $-0.05$          & $0.01$           & $-0.10$          \\
                      & $[-0.41;\ 0.42]$ & $[-0.21;\ 0.10]$ & $[-0.26;\ 0.28]$ & $[-0.28;\ 0.08]$ \\
TreatmentNo\_grass    & $-0.09$          &                  &                  &                  \\
                      & $[-0.75;\ 0.56]$ &                  &                  &                  \\
TreatmentNo\_shrub    &                  & $0.33^{*}$       & $-0.08$          & $0.21^{*}$       \\
                      &                  & $[0.01;\ 0.65]$  & $[-0.41;\ 0.25]$ & $[0.01;\ 0.40]$  \\
\hline
AIC                   & 284.97           & 160.41           & 337.77           & 253.15           \\
BIC                   & 298.99           & 172.62           & 352.65           & 268.46           \\
Log Likelihood        & -137.49          & -75.20           & -163.88          & -121.57          \\
Num. obs.             & 122              & 85               & 145              & 158              \\
Num. groups: quad     & 19               & 11               & 21               & 22               \\
Num. groups: year     & 9                & 9                & 9                & 9                \\
Var: quad (Intercept) & 0.43             & 0.00             & 0.02             & 0.00             \\
Var: year (Intercept) & 0.00             & 0.02             & 0.12             & 0.06             \\
Var: Residual         & 0.41             & 0.32             & 0.48             & 0.24             \\
\hline
\multicolumn{5}{l}{\scriptsize{$^*$ 0 outside the confidence interval}}
\end{tabular}
\label{table:coefficients}
%\end{center}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:32:17 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the \textit{A. tripartita} survival model} 
\label{ARTRsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & -0.0207 & 0.2261 & -0.4554 & -0.0246 & 0.4367 & -0.0321 & 0.0000 \\ 
  logarea & 0.7208 & 0.0575 & 0.6109 & 0.7195 & 0.8378 & 0.7171 & 0.0000 \\ 
  TreatmentNo\_grass & -1.3218 & 0.7343 & -2.7293 & -1.3340 & 0.1563 & -1.3591 & 0.0000 \\ 
  W.ARTR & -1.5214 & 0.2075 & -1.9391 & -1.5179 & -1.1237 & -1.5107 & 0.0000 \\ 
  W.HECO & -0.0694 & 0.0423 & -0.1485 & -0.0708 & 0.0178 & -0.0738 & 0.0000 \\ 
  W.POSE & -0.0049 & 0.0761 & -0.1469 & -0.0077 & 0.1525 & -0.0134 & 0.0000 \\ 
  W.PSSP & 0.0281 & 0.0275 & -0.0235 & 0.0271 & 0.0849 & 0.0252 & 0.0000 \\ 
  W.allcov & -0.0055 & 0.0083 & -0.0219 & -0.0054 & 0.0109 & -0.0054 & 0.0000 \\ 
  W.allpts & 0.1694 & 0.1579 & -0.1386 & 0.1686 & 0.4812 & 0.1672 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:33:00 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the \textit{H. comata} survival model} 
\label{HECOsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.4189 & 0.2115 & 0.9991 & 1.4190 & 1.8371 & 1.4192 & 0.0000 \\ 
  logarea & 1.2885 & 0.0813 & 1.1399 & 1.2843 & 1.4601 & 1.2758 & 0.0000 \\ 
  TreatmentNo\_shrub & 0.3958 & 0.4175 & -0.4193 & 0.3942 & 1.2190 & 0.3909 & 0.0000 \\ 
  W.ARTR & -0.0064 & 0.0028 & -0.0120 & -0.0064 & -0.0009 & -0.0064 & 0.0000 \\ 
  W.HECO & -0.6647 & 0.0673 & -0.7996 & -0.6638 & -0.5351 & -0.6619 & 0.0000 \\ 
  W.POSE & 0.0732 & 0.0534 & -0.0304 & 0.0728 & 0.1790 & 0.0720 & 0.0000 \\ 
  W.PSSP & 0.0190 & 0.0184 & -0.0167 & 0.0188 & 0.0554 & 0.0185 & 0.0000 \\ 
  W.allcov & 0.0021 & 0.0047 & -0.0072 & 0.0021 & 0.0113 & 0.0021 & 0.0000 \\ 
  W.allpts & -0.1185 & 0.0919 & -0.2998 & -0.1183 & 0.0610 & -0.1177 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:34:30 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the \textit{Poa secunda} survival model} 
\label{POSEsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.3162 & 0.1965 & 0.9242 & 1.3171 & 1.7024 & 1.3186 & 0.0000 \\ 
  logarea & 1.0585 & 0.0640 & 0.9365 & 1.0569 & 1.1902 & 1.0537 & 0.0000 \\ 
  TreatmentNo\_shrub & -0.2572 & 0.1723 & -0.5949 & -0.2573 & 0.0811 & -0.2576 & 0.0000 \\ 
  W.ARTR & 0.0001 & 0.0019 & -0.0036 & 0.0001 & 0.0038 & 0.0001 & 0.0000 \\ 
  W.HECO & -0.0153 & 0.0138 & -0.0424 & -0.0154 & 0.0119 & -0.0154 & 0.0000 \\ 
  W.POSE & -1.2632 & 0.0816 & -1.4260 & -1.2623 & -1.1053 & -1.2605 & 0.0000 \\ 
  W.PSSP & 0.0249 & 0.0122 & 0.0011 & 0.0249 & 0.0490 & 0.0247 & 0.0000 \\ 
  W.allcov & -0.0024 & 0.0026 & -0.0075 & -0.0024 & 0.0029 & -0.0024 & 0.0000 \\ 
  W.allpts & 0.0018 & 0.0524 & -0.1009 & 0.0018 & 0.1048 & 0.0017 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:35:48 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the \textit{P. spicata} survival model} 
\label{PSSPsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.2121 & 0.1699 & 0.8815 & 1.2103 & 1.5533 & 1.2066 & 0.0000 \\ 
  logarea & 1.5492 & 0.0976 & 1.3612 & 1.5474 & 1.7473 & 1.5437 & 0.0000 \\ 
  TreatmentNo\_shrub & -0.1885 & 0.2208 & -0.6237 & -0.1879 & 0.2431 & -0.1867 & 0.0000 \\ 
  W.ARTR & 0.0101 & 0.0021 & 0.0061 & 0.0101 & 0.0142 & 0.0101 & 0.0000 \\ 
  W.HECO & 0.0046 & 0.0185 & -0.0319 & 0.0046 & 0.0406 & 0.0047 & 0.0000 \\ 
  W.POSE & 0.0245 & 0.0384 & -0.0502 & 0.0242 & 0.1007 & 0.0236 & 0.0000 \\ 
  W.PSSP & -0.4430 & 0.0291 & -0.5010 & -0.4427 & -0.3867 & -0.4420 & 0.0000 \\ 
  W.allcov & 0.0100 & 0.0028 & 0.0046 & 0.0100 & 0.0155 & 0.0100 & 0.0000 \\ 
  W.allpts & 0.0943 & 0.0605 & -0.0246 & 0.0944 & 0.2129 & 0.0944 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:36:00 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the \textit{A. tripartita} growth model} 
\label{ARTRgrowth}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 0.7378 & 0.2437 & 0.2591 & 0.7378 & 1.2161 & 0.7377 & 0.0000 \\ 
  logarea.t0 & 0.8663 & 0.0393 & 0.7892 & 0.8663 & 0.9434 & 0.8663 & 0.0000 \\ 
  TreatmentNo\_grass & 0.1037 & 0.1436 & -0.1783 & 0.1037 & 0.3854 & 0.1037 & 0.0000 \\ 
  W.ARTR & -0.0364 & 0.0892 & -0.2116 & -0.0364 & 0.1387 & -0.0364 & 0.0000 \\ 
  W.HECO & 0.0033 & 0.0136 & -0.0233 & 0.0033 & 0.0299 & 0.0033 & 0.0000 \\ 
  W.POSE & -0.0573 & 0.0268 & -0.1100 & -0.0573 & -0.0047 & -0.0573 & 0.0000 \\ 
  W.PSSP & 0.0062 & 0.0084 & -0.0102 & 0.0062 & 0.0226 & 0.0062 & 0.0000 \\ 
  W.allcov & -0.0055 & 0.0030 & -0.0114 & -0.0055 & 0.0004 & -0.0055 & 0.0000 \\ 
  W.allpts & 0.0029 & 0.0477 & -0.0908 & 0.0029 & 0.0965 & 0.0029 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:36:24 2016
\begin{table}[ht]
\centering
\caption{Summary of fixed effects for the \textit{H. comata} growth model} 
\label{HECOgrowth}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 0.3782 & 0.0797 & 0.2212 & 0.3783 & 0.5343 & 0.3784 & 0.0000 \\ 
  logarea.t0 & 0.8272 & 0.0205 & 0.7869 & 0.8272 & 0.8674 & 0.8272 & 0.0000 \\ 
  TreatmentNo\_shrub & -0.0162 & 0.1336 & -0.2785 & -0.0162 & 0.2460 & -0.0162 & 0.0000 \\ 
  W.ARTR & -0.0031 & 0.0010 & -0.0051 & -0.0031 & -0.0010 & -0.0031 & 0.0000 \\ 
  W.HECO & -0.0479 & 0.0195 & -0.0862 & -0.0479 & -0.0098 & -0.0479 & 0.0000 \\ 
  W.POSE & 0.0242 & 0.0162 & -0.0077 & 0.0242 & 0.0560 & 0.0242 & 0.0000 \\ 
  W.PSSP & -0.0151 & 0.0065 & -0.0279 & -0.0151 & -0.0023 & -0.0151 & 0.0000 \\ 
  W.allcov & -0.0038 & 0.0019 & -0.0076 & -0.0038 & 0.0000 & -0.0038 & 0.0000 \\ 
  W.allpts & -0.0513 & 0.0380 & -0.1258 & -0.0513 & 0.0232 & -0.0513 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}
